Don't Mind the Gap: Bridging Network-wide Objectives and Device-level Configurations

### 2016-06-22 paper/v2/paper-edited.tex
more edits

     pol     &::=& p_1, \dots, p_n & \textit{policies} \\
     p       &::=& t \hspace{.3em} \Path \hspace{.3em} r_1 \Prefer \dots \Prefer r_m \BNFALT cc & \textit{constraints} \\
     x       &::=& d.d.d.d/d & \textit{prefix} \\
     t       &::=& \True & \textit{true} \\
         &\BNFALT& \NOT t & \textit{negation} \\
         &\BNFALT& t_1 \OR{} t_2 & \textit{disjunction} \\
         &\BNFALT& t_1 \AND{} t_2 & \textit{conjunction} \\
         &\BNFALT& \textit{prefix} = x & \textit{prefix test} \\
         &\BNFALT& \textit{comm} = d & \textit{community test} \\
     r       &::=& l & \textit{location} \\
         &\BNFALT& \emptyset & \textit{empty set} \\
         &\BNFALT& \In & \textit{internal loc} \\
         &\BNFALT& \Out & \textit{external loc} \\
         &\BNFALT& r_1 \cup r_2 & \textit{union} \\
         &\BNFALT& r_1 \cap r_2 & \textit{intersection} \\
         &\BNFALT& r_1 \cdot r_2 & \textit{concatenation} \\
         &\BNFALT& \NOT r & \textit{path negation} \\
         &\BNFALT& r^* & \textit{iteration} \\
     lp       &::=& r_1 \rightarrow r_2 & \textit{link pairs} \\
     cc     &::=& agg(x, lp) \BNFALT tag(d, t, lp) & \textit{control constraints} \\
  \end{array} \]

@@

\sysname also supports constraints on the control-plane behavior of BGP. For example, prefix aggregation is an important optimization to reduce routing table size. A constraint of the form $agg(x,lp)$ tells the compiler to perform aggregation for prefix $x$ across all links described by $lp$. It is also often useful to be able to add community tags to exported routes in BGP (\EG, to communicate non-standard information to peers). A constraint of the form $tag(d,t,lp)$ adds community tag $d$ for any prefixes matching $t$ across links $lp$.
%% Aggregation, for example, from internal to external locations, is specified using the same regular syntax as before:
%% $$\Agg(128.17.0.0/16, \In \Link \Out)$$
%% where the expression $\In \Link Out$ refers to control messages flowing from any internal to any external location.
We list only the route aggregation and community tagging constraints in Figure~\ref{fig:rir-syntax}, but \sysname also supports other constraints such as limiting the maximum number of routes allowed between ASes, or enabling BGP multipath.

@@

We use graph dominators~\cite{tarjan-dominance} as a relatively cheap approximation for removing many nodes and edges in the PGIR that are never on any \emph{simple} (loop free) path between the start and end nodes.
In the PGIR, a node $m$ dominates a node $n$ if $m$ appears on every path leading from the start node to $n$. Similarly, a node $m$ post-dominates a node $n$ in the PGIR if $m$ appears on every path from $n$ to the end node.
We can safely remove nodes and edges in the PGIR when any of the following conditions hold, where we have $m$, $m'$ and $n$, $n'$ such that $m \approx m'$ and $n \approx n'$.
\begin{itemize}
\setlength{\itemsep}{1pt}
\setlength{\parskip}{0pt}
\setlength{\parsep}{0pt}
\item Remove $m$ if it is not reachable from the start node
\item Remove $m$ if it can not reach the end node
\item Remove $m$ if it is (post-)dominated by some $m'$
\item Remove edge ($m$, $n$) if some $m'$ post-dominates $n$
\item Remove edge ($m$, $n$) if some $n'$ dominates $m$
\end{itemize}
For example, node $(W,1,1)$ in Figure~\ref{fig:example-compilation} is removed because every path to the end node must always go through node $(W,-,4)$. That is, node $(W,1,1)$ is post-dominated by node $(W,-,4)$ and both are shadows of topology location $W$.
Similarly, the edge from $(C,3,2)$ to $(D,-,2)$ is removed since node $(C,-,2)$ post-dominates $(D,-,2)$, and shadows the topology location $C$.

@@

The problem of determining local preferences for each router is reflected in the structure of the PGIR. Whenever a given router appears as multiple shadow nodes in the PGIR, the compiler must decide which shadow to prefer.
%
In the example from Figure~\ref{fig:example-compilation}, the topology node $C$ can receive an advertisement from $E$ in shadow $(C,-,2)$ or from $D$ in shadow $(C,3,2)$.
%
%
The compiler must determine a \emph{total ordering} of shadow nodes for each router, which reflects the relative preference of advertisements received in each shadow and should be consistent with path ranks in routing policy.
%
For example, if $C$'s shadow $(C,3,2)$ is preferred to $(C,-,2)$, written as $(C,3,2) \leq_{lp} (C,-,2)$,  $C$ will prefer advertisements from $(D,2,2)$ over $(E,-,2)$. $D$ and $E$ tag their advertisements to let $C$ know which shadow sent the advertisement. \$\ref{sec:abgp} discusses tagging in detail.

@@

\para{Regret-free Preferences}
To order PGIR shadows ($\leq_{lp}$) for each router in a way that is both sound and complete under all failures, we introduce the notion of \emph{regret-free} preferences, motivated by the observations from the example in Figure~\ref{fig:unimplementable}. A router (location) $l$ has a regret-free preference for a set of advertisements $A$ over $B$ if, whenever $l$ selects an advertisement to destination $d$ from $A$ over another from $B$, there is always some policy-compliant path to $d$ that is at least as good ($\leq_{rank}$ for the final node along the path) as any possible path (not necessily from $l$) to $d$ if $l$ had selected an advertisement from $B$ instead. In other words, the preference of $A$ over $B$ at $l$ is regret-free if $l$ is never (under any failure) worse-off by choosing an advertisement from $A$ when available. The notion of regret-free preferences can be lifted to PGIR nodes by considering the set of advertisements available to each node.

@@

On the other hand, this would not have been a problem if paths ending at $A_1$ had a lower rank than those ending at $A_2$ or $A_3$. For example, if paths ending at $A_1$ had rank 1, then any time $A_2$ is unusable due to a loop with $A_1$, it ultimately does not matter since $A_1$ is preferred anyway. 
%
In fact, checking if we are never worse off using $A_1$ instead of $A_2$ corresponds exactly with determining if $A$ has a \textup{Regret-Free} preference for $A_1$ over $A_2$.
%
More specifically, the compiler checks that, any time there are two shadows $N_1$ and $N_2$ for the same topology location, where $N_1$ appears ``above'' (\IE, can reach) $N_2$ in the PGIR, then $N_1$ must be strictly preferred to $N_2$ (\IE, $N_1 <_{lp} N_2$).

@@

Instead, we adopt the following simple strategy: $i)$ pick a random path in the PGIR between the prefix origin and aggregation point, $ii)$ remove all similar edges in the PGIR for each topological edges along the chosen path, and $iii)$ repeat until no such path exists.
Because each path chosen is both policy compliant and edge disjoint (due to ii), the number of paths that we are able to remove lower bounds the number of failures required to disconnect the prefix from its aggregate, subject to the policy constraints.

@@

Constructing automata for extended regular expressions (\IE, regular expressions with negation and intersection operations) is known to have high complexity~\cite{regex-complexity}. The \sysname compiler uses regular expressions derivatives~\cite{regex-derivatives} with character classes to construct deterministic automata for extended regular expressions efficiently. Since regular expressions are defined over a finite alphabet, and since much of the AS topology is unknown, we set the alphabet to include all uniquely referenced external ASes in the policy. To model the unknown external AS topology beyond immediate peers, we include a special topology node to represent any unknown location.
%
Rather than construct the product graph in full, our implementation prevents exploring parts of the graph during construction when no automata has a reachable accepting state.

@@

When computing local preferences and ensuring failure safety, as described in \S\ref{sec:compilation}, the compiler performs memoization of the \textup{Regret-Free} function. That is, whenever for two states $N_1$ and $N_2$ we compute $\textup{Regret-Free}(G, N_1, N_2)$, if the function evaluates to $true$, then each of the intermediate related states $m$ and $n$ must also satisfy $\textup{Regret-Free}(G, m, n)$. Memoizing these states dramatically reduces the amount of work performed to find preferences in the common case.

@@

The naive code generation algorithm described in \S\ref{sec:compilation} is extremely memory inefficient since it generates a separate match-export pair for every unique in-edge/out-edge pair for every node in the product graph before minimization. Our implementation performs partial minimization during generation by recognizing common cases such as when there is no restriction on exporting to or importing from neighbors.

@@

Even when programming the network centrally, it is possible for operators to make mistakes and write incorrect policies.
\sysname includes many analyses to identify common mistakes at compile time. A subset includes: (i) a preference analysis to determine when backup paths/preferences will never be used, (ii) a reachability analysis to check if locations that should be reachable according to the policy are not reachable after combining both the topology and policy, (iii) an anycast analysis to find instances where the operator might accidentally anycast a prefix (\IE, originates the prefix from multiple locations), (iv) an aggregate analysis to find when aggregates do not summarize any specific prefix.

@@

We apply \sysname on real policies for backbone and datacenter networks. Our main goals are to evaluate if its language is expressive enough for real-world policies and the time the compiler takes to generate router configurations.

@@

We found that we could translate all network policies to \sysname. We verified with the operators that our translation preserved intended semantics.\footnote{Not intended as a scientific test, but we also asked the two operators if they would find it easy to express their policies in \sysname. The datacenter operator said that he found the language intuitive. The backbone operator said that formalizing the policy in \sysname seemed equally easy or difficult as formalizing in RPSL~\cite{RFC2622}, but he appreciated that he would have to do it only once for the whole network (not per-router) and did not have to manually compute various local preferences, import-export filters, and MEDs.} We further found that the datacenter policies were correctly translated. For the backbone network, the operator mentioned an additional policy not present in the English document, which we added later.

@@

Total compilation for the largest datacenter is less than 9 minutes total. Unlike the datacenter policy, the number of prefixes for the backbone policy remains relatively fixed as the topology size increases. Compilation for the largest backbone network, takes less than 3 minutes total. The inclusion of both more preferences and more neighboring ASes in the backbone policy increases the size of the resulting PGIR, which in turn cleads to PGIR construction and minimization taking proportionally more time.

>>>>

     pol     &::=& p_1, \dots, p_n & \textit{policies} \\
     p       &::=& t \hspace{.3em} \Path \hspace{.3em} r_1 \Prefer \dots \Prefer r_m \BNFALT cc & \textit{constraints} \\
     x       &::=& d.d.d.d/d & \textit{prefix} \\
     t       &::=& \True & \textit{true} \\
         &\BNFALT& \NOT t & \textit{negation} \\
         &\BNFALT& t_1 \OR{} t_2 & \textit{disjunction} \\
         &\BNFALT& t_1 \AND{} t_2 & \textit{conjunction} \\
         &\BNFALT& \textit{prefix} = x & \textit{prefix test} \\
         &\BNFALT& \textit{comm} = d & \textit{community test} \\
     r       &::=& l & \textit{location} \\
         &\BNFALT& \emptyset & \textit{empty set} \\
         &\BNFALT& \In & \textit{internal loc} \\
         &\BNFALT& \Out & \textit{external loc} \\
         &\BNFALT& r_1 \cup r_2 & \textit{union} \\
         &\BNFALT& r_1 \cap r_2 & \textit{intersection} \\
         &\BNFALT& r_1 \cdot r_2 & \textit{concatenation} \\
         &\BNFALT& \NOT r & \textit{path negation} \\
         &\BNFALT& r^* & \textit{iteration} \\
     ln       &::=& r_1 \rightarrow r_2 & \textit{links} \\
     cc     &::=& agg(x, ln) \BNFALT tag(d, t, ln) & \textit{control constraints} \\
  \end{array} \]

@@

\sysname also supports constraints on the control-plane behavior of BGP. For example, prefix aggregation is an important optimization to reduce routing table size. A constraint of the form $agg(x,ln)$ tells the compiler to perform aggregation for prefix $x$ across all links described by $ln$. It is also often useful to be able to add community tags to exported routes in BGP (\EG, to communicate non-standard information to peers). A constraint of the form $tag(d,t,ln)$ adds community tag $d$ for any prefixes matching $t$ across links $ln$.
%% Aggregation, for example, from internal to external locations, is specified using the same regular syntax as before:
%% $$\Agg(128.17.0.0/16, \In \Link \Out)$$
%% where the expression $\In \Link Out$ refers to control messages flowing from any internal to any external location.
We list only the route aggregation and community tagging constraints in Figure~\ref{fig:rir-syntax}, but \sysname also supports other constraints such as limiting the maximum number of routes allowed between ASes, or enabling BGP multipath.

@@

We use graph dominators~\cite{tarjan-dominance} as a relatively cheap approximation for removing many nodes and edges in the PGIR that are never on any \emph{simple} (loop free) path between the start and end nodes.
In the PGIR, a node $m$ dominates a node $n$ if $m$ appears on every path leading from the start node to $n$. Similarly, a node $m$ post-dominates a node $n$ in the PGIR if $m$ appears on every path from $n$ to the end node.
We can safely remove nodes and edges in the PGIR when any of the following conditions hold, where we have $m$, $m'$ and $n$, $n'$ such that $m \approx m'$ and $n \approx n'$.
\begin{itemize}
\setlength{\itemsep}{1pt}
\setlength{\parskip}{0pt}
\setlength{\parsep}{0pt}
\item Remove $m$ if it is not reachable from the start node
\item Remove $m$ if it can not reach the end node
\item Remove $m$ if it is (post-)dominated by some $m'$
\item Remove edge ($m$, $n$) if some $m'$ post-dominates $n$
\item Remove edge ($m$, $n$) if some $n'$ dominates $m$
\end{itemize}
For example, node $(W,1,1)$ in Figure~\ref{fig:example-compilation} is removed because every path to the end node must always go through node $(W,-,4)$. That is, node $(W,1,1)$ is post-dominated by node $(W,-,4)$ and both are shadows of topology location $W$.
Similarly, the edge from $(C,3,2)$ to $(D,-,2)$ is removed since node $(C,-,2)$ post-dominates $(D,-,2)$.

@@

The problem of determining local preferences for each router is reflected in the structure of the PGIR. Whenever a given router appears as multiple shadow nodes in the PGIR, the compiler must decide which shadow to prefer.
%
In the example from Figure~\ref{fig:example-compilation}, the topology node $C$ can receive an advertisement from $E$ in shadow $(C,-,2)$ or from $D$ in shadow $(C,3,2)$.
%
%
The compiler must determine a \emph{total ordering} of shadow nodes for each router, which reflects the relative preference of advertisements received in each shadow and should be consistent with path ranks in routing policy.
%
For example, if $C$'s shadow $(C,3,2)$ can be preferred to $(C,-,2)$, written as $(C,3,2) \leq_{lp} (C,-,2)$, $C$ can prefer advertisements from $(D,2,2)$ over $(E,-,2)$. $D$ and $E$ tag their advertisements to let $C$ know which shadow sent the advertisement. \$\ref{sec:abgp} discusses tagging in detail.

@@

\para{Regret-free Preferences}
To order PGIR shadows ($\leq_{lp}$) for each router in a way that is policy-compliant under all failures, we introduce the notion of \emph{regret-free} preferences, motivated by the observations from the example in Figure~\ref{fig:unimplementable}. A router (location) $l$ has a regret-free preference for a set of advertisements $A$ over $B$ if, whenever $l$ selects an advertisement to destination $d$ from $A$ over another from $B$, there is always some policy-compliant path to $d$ that is at least as good ($\leq_{rank}$ for the final node along the path) as any possible path (not necessarily from $l$) to $d$ if $l$ had selected an advertisement from $B$ instead. In other words, the preference of $A$ over $B$ at $l$ is regret-free if $l$ is never (under any failure) worse-off by choosing an advertisement from $A$ when available. The notion of regret-free preferences can be lifted to PGIR nodes by considering the set of advertisements available to each node.

@@

On the other hand, this would not have been a problem if paths ending at $A_1$ had a lower rank than those ending at $A_2$ or $A_3$. For example, if paths ending at $A_1$ had rank 1, then any time $A_2$ is unusable due to a loop with $A_1$, it ultimately does not matter since $A_1$ is preferred anyway. 
%
In fact, checking if we are never worse off using $A_1$ instead of $A_2$ corresponds exactly with determining if $A$ has a regret-free preference for $A_1$ over $A_2$.
%
More specifically, the compiler checks that, any time there are two shadows $N_1$ and $N_2$ for the same topology location, where $N_1$ appears ``above'' (\IE, can reach) $N_2$ in the PGIR, then $N_1$ must be strictly preferred to $N_2$ (\IE, $N_1 <_{lp} N_2$).

@@

Instead, we adopt the following simple strategy: $i)$ pick a random path in the PGIR between the prefix's origin and aggregation point, $ii)$ remove all similar edges in the PGIR for each topology edge along the chosen path, and $iii)$ repeat until no such path exists.
Because each path chosen is both policy compliant and edge disjoint (due to ii), the number of paths that we are able to remove lower bounds the number of failures required to disconnect the prefix from its aggregate, subject to the policy constraints.

@@

Constructing automata for extended regular expressions (\IE, regular expressions with negation and intersection operations) is known to have high complexity~\cite{regex-complexity}. The \sysname compiler uses regular expression derivatives~\cite{regex-derivatives} with character classes to construct deterministic automata for extended regular expressions efficiently. Since regular expressions are defined over a finite alphabet, and since much of the AS topology is unknown, we set the alphabet to include all uniquely referenced external ASes in the policy. To model the unknown external AS topology beyond immediate peers, we include a special topology node to represent any unknown location.
%
Rather than construct the product graph in full, our implementation prevents exploring parts of the graph during construction when no automata has a reachable accepting state.

@@

When computing local preferences and ensuring failure safety, as described in \S\ref{sec:compilation}, the compiler performs memoization of the \textup{Regret-Free} function. That is, whenever for two states $N_1$ and $N_2$ we compute $\textup{Regret-Free}(G, N_1, N_2)$ and the function evaluates to $true$, then each of the intermediate related states $m$ and $n$ must also satisfy $\textup{Regret-Free}(G, m, n)$. Memoizing these states dramatically reduces the amount of work performed to find preferences in the common case.

@@

The naive code generation algorithm described in \S\ref{sec:abgp} is extremely memory inefficient since it generates a separate match-export pair for every unique in-edge/out-edge pair for every node in the product graph before minimization. Our implementation performs partial minimization during generation by recognizing common cases such as when there is no restriction on exporting to or importing from neighbors.

@@

Even when programming the network centrally, it is possible for operators to make mistakes and write incorrect policies.
\sysname includes many analyses to identify common mistakes at compile time. A subset includes: (i) a preference analysis to determine when backup paths/preferences will never be used, (ii) a reachability analysis to check if locations that should be reachable according to the policy are not reachable after combining both the topology and policy, (iii) an anycast analysis to find instances where the operator might accidentally anycast a prefix (\IE, originates the prefix from multiple locations), (iv) an aggregate analysis to find aggregates that do not summarize any specific prefix.

@@

We apply \sysname on real policies for backbone and datacenter networks. Our main goals are to evaluate if its language is expressive enough for real-world policies, the time the compiler takes to generate router configurations, and the size of the resulting configurations.

@@

We found that we could translate all network policies to \sysname. We verified with the operators that our translation preserved intended semantics.\footnote{Not intended as a scientific test, but we also asked the two operators if they would find it easy to express their policies in \sysname. The datacenter operator said that he found the language intuitive. The backbone operator said that formalizing the policy in \sysname seemed equally easy or difficult as formalizing in RPSL~\cite{RFC2622}, but he appreciated that he would have to do it only once for the whole network (not per-router) and did not have to manually compute various local preferences, import-export filters, and MEDs.} We further found that the datacenter policies were correctly translated. For the backbone network, the operator mentioned an additional policy not present in the English document, which we added later.
%
For both the datacenter and backbone networks, the \sysname compiler was able to guarantee policy-compliance under all possible failure scenarios.

@@

Total compilation for the largest datacenter is less than 9 minutes total. Unlike the datacenter policy, the number of prefixes for the backbone policy remains relatively fixed as the topology size increases. Compilation for the largest backbone network, takes less than 3 minutes total. The inclusion of both more preferences and more neighboring ASes in the backbone policy increases the size of the resulting PGIR, which in turn leads to PGIR construction and minimization taking proportionally more time.
